---
title: "SARIMA"
author: "Yeseul_Sung"
date: "7/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
setwd("/Users/seong-yeseul/Downloads")
library(data.table)
library(forecast)
library(dplyr)
library(tidyverse)
library(magrittr)
source("TS-library.R")
source("TS-yslib.R")
```

# Cluster 0

## Cluster Sum

```{r message=FALSE, warning=FALSE}
data = fread("v23gb2_cluster0.csv")[, 4:56] %>% apply(2, sum)
data = ts(data, frequency = 12, start = c(2017, 01))
plot.ts(data) #seasonality? linear or poly trend?

acf(data, lag = 50) #till lag 10, lag 37
pacf(data, lag = 50) #1
plot.ts(diff(data))
plot.ts(diff(diff(data))) #VST needed maybe
```


```{r message=FALSE, warning=FALSE}
library(MASS)
fit = boxcox(data ~ 1)
lambda = fit$x[which.max(fit$y)]; lambda

dat = log(data)
dat2 = data ^(-2)
par(mfrow = c(1, 3))
plot.ts(data)
plot.ts(dat)
plot.ts(dat2)

qqnorm(data); qqline(data)
qqnorm(dat); qqline(dat)
qqnorm(dat2);qqline(dat2)

acf(data, lag = 50)
acf(dat, lag = 50)
acf(dat2, lag = 50)

pacf(data, lag = 50)
pacf(dat, lag = 50)
pacf(dat2, lag = 50)
```

```{r message=FALSE, warning=FALSE}
#auto.arima
auto.arima(data) #(0, 1, 2) 0.8977 log likelihood=-70.1 AIC=146.21
auto.arima(dat) #(0, 1, 0) 0.006872:log likelihood=55.7 AIC=-109.4
auto.arima(dat2) #(0, 1, 0)1.546e-06:log likelihood=274.09 AIC=-546.18 
par(mfrow = c(1, 1))
```


```{r echo=TRUE, message=FALSE, warning=FALSE}
#original data
out = compare_stats(data, 2, 2, 2, 1, 1, 1, period = 12);out[1:10, ] 
```


```{r message=FALSE, warning=FALSE}
out_error_plot(data, 37, a_order = c(0, 1, 2)) #2.167863
out_error_plot(data, 37, a_order = c(2, 1, 0), s_order = c(0, 1, 1), period = 12)#4.372653
out_error_plot(data, 37, a_order = c(0, 1, 2), s_order = c(0, 1, 1), period = 12)#4.850387
out_error_plot(data, 37, a_order = c(0, 1, 1), s_order = c(0, 1, 1), period = 12)#4.547179
out_error_plot(data, 37, a_order = c(1, 1, 1), s_order = c(0, 1, 1), period = 12)#4.678489
out_error_plot(data, 37, a_order = c(2, 1, 1), s_order = c(0, 1, 1), period = 12)#4.672863

```

in terms of AIC/BIC/sigma2/log likelihood, fit2 > 3 > 4 > 5 > 6 > 1
in terms of out of sample MSE fit1 > 2 > 4 > 6 > 5 > 3

```{r message=FALSE, warning=FALSE}
fit1 = arima(data, order = c(0, 1, 2), include.mean = F)
fit2 = arima(data, order = c(2, 1, 0), 
             seasonal = list(order = c(0, 1, 1), period = 12), include.mean = F)
fit3 = arima(data, order = c(0, 1, 2), 
             seasonal = list(order = c(0, 1, 1), period = 12), include.mean = F)
fit4 = arima(data, order = c(0, 1, 1), 
             seasonal = list(order = c(0, 1, 1), period = 12), include.mean = F)
fit5 = arima(data, order = c(1, 1, 1), 
             seasonal = list(order = c(0, 1, 1), period = 12), include.mean = F)
fit6 = arima(data, order = c(2, 1, 1), 
             seasonal = list(order = c(0, 1, 1), period = 12), include.mean = F)

#to test if coefficients are away from 0 or not
test_coef(fit1)
test_coef(fit2)
test_coef(fit3)
test_coef(fit4)
test_coef(fit5)
test_coef(fit6)
```

in terms of coefficients test, fit4> fit5 > fit1 > fit2 > fit3 > fit6

```{r message=FALSE, warning=FALSE}
test(residuals(fit1)) #0 rejected, acf/pacf/resi/qq ok
test(residuals(fit2)) #1 rejected, resi & qq seems to be unstable
test(residuals(fit3)) #1 rejected, same as fit2
test(residuals(fit4)) #2 rejected, same
test(residuals(fit5)) #1 rejected, same
test(residuals(fit6)) #1 rejected, same
```

in terms of residuals, fit1 > fit2, 4, 5, 6 > fit3

our final model is fit 2, 4

```{r message=FALSE, warning=FALSE}
plot(forecast::forecast(fit2))
plot(forecast::forecast(fit4))
```

by the same procedure as above, the best model in log transformed data is as below:

```{r message=FALSE, warning=FALSE}
fit23 = arima(dat, order = c(1, 1, 1), 
      seasonal = list(order = c(0, 1, 0), period = 12), 
      include.mean = F)
out_error_plot(dat, 37, a_order = c(1, 1, 1), 
               s_order = c(0, 1, 0), period = 12) #0.02959237
test_coef(fit23) #all T
test(residuals(fit23)) #0 rejected, lag 12
plot(forecast::forecast(fit23))
```

to compare with prophet, out of sample prediction error in original scale is as below:

```{r message=FALSE, warning=FALSE}
n = length(dat)
xhat = numeric(n -37)
for (i in 0:(n - 38)){
  fit = arima(dat[1:(37 + i)], order = c(1, 1, 1), 
              seasonal = list(order = c(0, 1, 0), period = 12),
              include.mean = F)
  xhat[i+1] = exp(forecast::forecast(fit, h = 1)$mean)
}
mean((data[38:53] - xhat)^2) #original scale out of sample MSE
```

to compare with prophet, mape in original scale is as below :
```{r message=FALSE, warning=FALSE}
#mape
n = length(dat)
fit = arima(dat[1:37], order = c(1, 1, 1), seasonal = list(order = c(0, 1, 0), period =12), include.mean = F)
xhat = exp(forecast::forecast(fit, h = 16)$mean)
mape = (1 - mean(abs((xhat - data[38:53])/data[38:53])))*100
mape
```

## Cluster Median

```{r message=FALSE, warning=FALSE}
data = fread("arima_median.csv")[1,-1] %>% apply(2, sum)
data = ts(data, frequency  = 12, start = c(2017, 01))
plot.ts(data)
acf(data, lag = 50)
pacf(data, lag = 50)
plot.ts(diff(data))
plot.ts(diff(diff(data))) #VST needed
library(MASS)
fit = boxcox(data ~ 1)
lambda = fit$x[which.max(fit$y)]; lambda
dat = log(data)
```

cluster0 median final model : (1, 1, 1)(0, 1, 0)[12]

```{r message=FALSE, warning=FALSE}
fit23 = arima(dat, order = c(1, 1, 1), seasonal = list(order = c(0, 1, 0), period = 12),
             include.mean = F)
out_error_plot(dat, 37, a_order = c(1, 1, 1), s_order = c(0, 1, 0), period = 12) #0.0343151
test_coef(fit23) #2/3 **
test(residuals(fit23)) #ok
plot(forecast::forecast(fit23))
```

```{r message=FALSE, warning=FALSE}
n = length(dat)
xhat = numeric(n -37)
for (i in 0:(n - 38)){
  fit = arima(dat[1:(37 + i)], order = c(1, 1, 1), 
              seasonal = list(order = c(0, 1, 0), period = 12),
              include.mean = F)
  xhat[i+1] = exp(forecast::forecast(fit, h = 1)$mean)
}
mean((data[38:53] - xhat)^2) #original scale out of sample MSE
```

to compare with prophet, mape in original scale is as below :
```{r message=FALSE, warning=FALSE}
#mape
n = length(dat)
fit = arima(dat[1:37], order = c(1, 1, 1), seasonal = list(order = c(0, 1, 0), period =12), include.mean = F)
xhat = exp(forecast::forecast(fit, h = 16)$mean)
mape = (1 - mean(abs((xhat - data[38:53])/data[38:53])))*100
mape
```

#Cluster 1

## Cluster sum

```{r message=FALSE, warning=FALSE}
data = fread("v23gb2_cluster1.csv")[, 4:56] %>% apply(2, sum)
data = ts(data, frequency = 12, start = c(2017, 01))
plot.ts(data) #seasonality? poly trend?
acf(data, lag = 50) #~10, 37
pacf(data, lag = 50) #1
plot.ts(diff(data))
plot.ts(diff(diff(data)))
library(MASS)
fit = boxcox(data ~ 1)
lambda = fit$x[which.max(fit$y)]; lambda
dat = log(data)
dat2 = data^(-2)
auto.arima(data) #(0,1,1)(1,0,0)[12] 0.1697:  log likelihood=-27.06 AIC=60.13
auto.arima(dat) #0.005294:  log likelihood=62.99 AIC=-119.99 
auto.arima(dat2) #2.275e-05:  log likelihood=204.52 AIC=-403.04
```

```{r message=FALSE, warning=FALSE}
out = compare_stats(data, 1,1, 1, 1, 1, 1, period = 12);out
fit1 = arima(data, order = c(0, 1, 1), 
             seasonal = list(order = c(1, 0, 0), period = 12), 
             include.mean = F) #60.12625  65.97999 0.1631558  -27.06313
fit2 = arima(data, order = c(0, 1, 1), 
             seasonal = list(order = c(0, 1, 1), period = 12), 
             include.mean = F) #58.25701  63.32365 0.1386916  -26.12850
fit3 = arima(data, order = c(1, 1, 1), 
             seasonal = list(order = c(0, 1, 1), period = 12), 
             include.mean = F) #59.96654  66.72206 0.1377942  -25.98327
fit4 = arima(data, order = c(1, 0, 1),
             seasonal = list(order = c(0, 1, 1), period = 12),
             include.mean = F) #60.01738  66.87166 0.1343796  -26.00869
#fit2 > fit3 > fit4 > fit1
out_error_plot(data, 37, a_order = c(0, 1, 1), s_order = c(1, 0, 0), period = 12) #0.4071355
out_error_plot(data, 37, a_order = c(0, 1, 1), s_order = c(0, 1, 1), period = 12) #0.6578255
out_error_plot(data, 37, a_order = c(1, 1, 1), s_order = c(0, 1, 1), period = 12) #0.6757334
#fit1 > fit2 > fit3 > fit4
test_coef(fit1)  #sar1 F
test_coef(fit2) #sma F
test_coef(fit3) #ar1, sma1 F
test_coef(fit4) #sma F
#fit4 > fit1 = fit2 > fit3
```
```{r message=FALSE, warning=FALSE}
test(residuals(fit1)) #0 rejected, everythings ok
test(residuals(fit2)) #2 rejected, residual&qq seems unstable, acf lag 5
test(residuals(fit3)) #1 rejected, residual & qq nono, acf lag 5
test(residuals(fit4)) #0 rejected, resi qq nono
```


final model in original dataset is fit1
```{r message=FALSE, warning=FALSE}
plot(forecast::forecast(fit1))
```

by the same procedure as above, final model is SARIMA(1, 0, 2)(0, 0, 1)[12] with data(-2)

```{r message=FALSE, warning=FALSE}
fit6 = arima(dat2, order = c(1, 0, 2),
             seasonal = list(order = c(0, 0, 1), period = 12),
             include.mean = F) #-404.7524 -394.9009 2.104561e-05 207.3762
out_error_plot(dat2, 37, a_order = c(1, 0, 2), s_order = c(0, 0, 1), period = 12) # 3.594744e-05
test_coef(fit6) #ma1 ma2
test(residuals(fit6)) #0, "
plot(forecast::forecast(fit6))
```

to compare with prophet,

```{r message=FALSE, warning=FALSE}
n = length(dat2)
xhat = numeric(n -37)
for (i in 0:(n - 38)){
  fit = arima(dat[1:(37 + i)], order = c(1, 1, 1), 
              seasonal = list(order = c(0, 1, 0), period = 12),
              include.mean = F)
  fore = forecast::forecast(fit, h = 1)$mean
  if (fore > 0){
    xhat[i+1] = sqrt(fore)^(-1)
  } else{
    xhat[i + 1] = 0
  }
  
}
mean((data[38:53] - xhat)^2) #original scale out of sample MSE
```

to compare with prophet, mape in original scale is as below :
```{r message=FALSE, warning=FALSE}
#mape
n = length(dat2)
fit = arima(dat[1:37], order = c(1, 1, 1), seasonal = list(order = c(0, 1, 0), period =12), include.mean = F)
xhat = forecast::forecast(fit, h = 16)$mean
xhat = sqrt(ifelse(xhat > 0, xhat, 0))^(-1)
mape = (1 - mean(abs((xhat - data[38:53])/data[38:53])))*100
mape
```

## Cluster 1 : Median

```{r message=FALSE, warning=FALSE}
data = fread("arima_median.csv")[2, -1] %>% apply(2, sum)
data = ts(data, frequency = 12, start = c(2017, 01))
plot.ts(data)
plot.ts(diff(data))
plot.ts(diff(diff(data)))
library(MASS)
fit = boxcox(data ~ 1)
lambda = fit$x[which.max(fit$y)]; lambda
dat = log(data)
```

```{r message=FALSE, warning=FALSE}
fit4 = arima(data, order = c(0, 1, 1), seasonal = list(order = c(1, 0, 1), period = 12),
             include.mean = F)
out_error_plot(data, 37, a_order = c(0, 1, 1), s_order = c(1, 0, 1), period = 12) #8.868089e-11
test_coef(fit4)#2/3
test(residuals(fit4))#okkk
plot(forecast::forecast(fit4))
```

to compare with prophet, mape in original scale is as below :
```{r message=FALSE, warning=FALSE}
mspe_plot(data, 37, a_order = c(0, 1, 1), s_order = c(1, 0, 1), period = 12)
```
# Cluster 2

## Cluster 2 : Sum

```{r message=FALSE, warning=FALSE}
data = fread("v23gb2_cluster2.csv")[, 4:56] %>% apply(2, sum)
data = ts(data, frequency = 12, start = c(2017, 01))
plot.ts(data) #seasonality? poly trend?
acf(data, lag = 50) #~10
pacf(data, lag = 50) #1
plot.ts(diff(data))
plot.ts(diff(diff(data)))

auto.arima(data) #ARIMA(0,1,0) 0.7899: -67.65 AIC=137.31 

fit3 = arima(data, order = c(1, 1, 1),
             seasonal = list(order = c(0, 1, 1), period = 12), 
             include.mean = F) #21.9999 128.7555  0.6464910  -56.99997
out_error_plot(data, 37,a_order = c(1, 1, 1), s_order = c(0, 1, 1), period = 12) #3.269203

test_coef(fit3)
test(residuals(fit3)) #1 rejected
plot(forecast::forecast(fit3))
```

to compare w prophet, mape in original scale is as below:
```{r message=FALSE, warning=FALSE}
mspe_plot(data, 37, a_order = c(1, 1, 1), s_order = c(0, 1, 1), period = 12)
```


## Cluster 2 : Median data

```{r message=FALSE, warning=FALSE}
data = fread("arima_median.csv")[3,-1] %>% apply(2, sum)
data = ts(data, frequency  = 12, start = c(2017, 01))
plot.ts(data)
acf(data, lag = 50) #till lag 8
pacf(data, lag = 50) #lag 1
plot.ts(diff(data))
plot.ts(diff(diff(data))) #vst needed maybe
library(MASS)
fit = boxcox(data ~ 1)
lambda = fit$x[which.max(fit$y)]; lambda #lambda near 0
dat = log(data)
```

final model

```{r message=FALSE, warning=FALSE}
fit1 = arima(dat, order = c(1, 1, 1), season = list(order = c(1, 0, 0), period = 12),
             include.mean = F)
out_error_plot(dat, 37, a_order = c(1, 1, 1), s_order = c(1, 0, 0), period = 12) #0.01792916
test_coef(fit1)
test(residuals(fit1)) #ok
plot(forecast::forecast(fit1))
```


```{r message=FALSE, warning=FALSE}
n = length(dat)
xhat = numeric(n -37)
for (i in 0:(n - 38)){
  fit = arima(dat[1:(37 + i)], order = c(1, 1, 1), 
              seasonal = list(order = c(1, 0, 0), period = 12),
              include.mean = F)
  xhat[i+1] = exp(forecast::forecast(fit, h = 1)$mean)
}
mean((data[38:53] - xhat)^2) #original scale out of sample MSE
```

to compare with prophet, mape in original scale is as below :
```{r message=FALSE, warning=FALSE}
#mape
n = length(dat)
fit = arima(dat[1:37], order = c(1, 1, 1), seasonal = list(order = c(1, 0, 0), period =12), include.mean = F)
xhat = exp(forecast::forecast(fit, h = 16)$mean)
mape = (1 - mean(abs((xhat - data[38:53])/data[38:53])))*100
mape
```